---
title: "GLM and Logistic Regression with LASSO/Ridge Regularization"
author: "Tianze Hua"
date: "3/4/2022"
output: pdf_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, tidy.opts=list(width.cutoff=80), tidy=TRUE)
options(tinytex.verbose = TRUE)
```


```{r load package and dataset}
# load multiple packages with two lines
library(pacman)
p_load(ISLR, caret, tidyverse, gridExtra, pROC, psych, knitr,
       broom, gmodels, glmnet, Metrics)

# dataset used is College
attach(College)
md = College
```

#### Context

Description
Statistics for a large number of US Colleges from the 1995 issue of US News and World Report.

#### Format

A data frame with 777 observations on the following 18 variables.

**Private** A factor with levels No and Yes indicating private or public university

**Apps Number** of applications received

**Accept** Number of applications accepted

**Enroll** Number of new students enrolled

**Top10perc** Pct. new students from top 10% of H.S. class

**Top25perc** Pct. new students from top 25% of H.S. class

**F.Undergrad** Number of fulltime undergraduates

**P.Undergrad** Number of parttime undergraduates

**Outstate** Out-of-state tuition

**Room.Board** Room and board costs

**Books** Estimated book costs

**Personal** Estimated personal spending

**PhD** Pct. of faculty with Ph.D.’s

**Terminal** Pct. of faculty with terminal degree

**S.F.Ratio** Student/faculty ratio

**perc.alumni** Pct. alumni who donate

**Expend** Instructional expenditure per student

**Grad.Rate** Graduation rate



```{r descriptive statistics }
glimpse(md)

psych::describe(md, fast=TRUE) %>% slice(2:n()) %>% select(3:8) %>% kable()

# Frequency table for Private College and Public
table(factor(md$Private))

# set Private college to be the baseline
md$Private <- relevel(md$Private, 'No')

# descriptive statistics for continuous variables
continous = select_if(md, is.numeric)
summary(continous)
```

```{r Partition trainning set}
set.seed(16)
train_index <- createDataPartition(md$Private, p=0.75, list = FALSE, times = 1)
train_data <- md[train_index,]
test_data <- md[-train_index,]
```

```{r fit the logistic regression model}
model <- glm(
  Private ~ Personal + PhD + Grad.Rate + S.F.Ratio,
  data = md,
  family = binomial(link = 'logit'))

summary(model)

# convert coefficient log odds into odds
exp(coef(model)) %>% kable()
```

```{r train set prediction}
prob_train <- predict(model, newdata = train_data, type = 'response')
predict_result <- as.factor(ifelse(prob_train >= 0.5, 'Yes', 'No'))
```

```{r confusion matrix}
confusionMatrix(predict_result, train_data$Private, positive = 'Yes')
CrossTable(predict_result, train_data$Private)
```

```{r test set}
prob_test <- predict(model, newdata = test_data, type = 'response')
predict_result <- as.factor(ifelse(prob_test >= 0.5, 'Yes', 'No'))

confusionMatrix(predict_result, test_data$Private, positive = 'Yes')
CrossTable(predict_result, test_data$Private)
```

```{r ROC curve}
curve <- roc(train_data$Private, prob_train)
plot(curve, 
     col = 'Blue', ylab='Sensitivity - TP Rate', 
     xlab = 'Specificity - FP Rate', main="ROC Curve")

pROC::auc(curve)
auc
```

```{r}
set.seed(16)
train_index <- createDataPartition(md$Grad.Rate, p=0.8, list = FALSE, times = 1)
train <- md[train_index,]
test <- md[-train_index,]

train_x <- model.matrix(Grad.Rate ~., train)[,-1]
test_x <- model.matrix(Grad.Rate ~., test)[,-1]

train_y <- train$Grad.Rate
test_y <- test$Grad.Rate
```

```{r}
cv.lasso <- cv.glmnet(train_x, train_y, nfolds = 10)
plot(cv.lasso)

log(cv.lasso$lambda.min)
log(cv.lasso$lambda.1se)

cv.lasso$lambda.min
```

```{r}
# alpha = 1 for Lasso(L2)
# alpha = 0 for Ridge(L1)
model.min <- glmnet(train_x, train_y, alpha = 1, lambda = cv.lasso$lambda.min)
model.min
coef(model.min)

model.1se <- glmnet(train_x, train_y, alpha = 1, lambda = cv.lasso$lambda.1se)
model.1se
coef(model.1se)
```

```{r}
model1.min <- glmnet(train_x, train_y, alpha = 0, lambda = cv.lasso$lambda.min)
model1.min
coef(model1.min)

model1.1se <- glmnet(train_x, train_y, alpha = 0, lambda = cv.lasso$lambda.1se)
model1.1se
coef(model1.1se)
```


```{r}
ols <- lm(Grad.Rate ~., data = train)
summary(ols)
```

```{r}
preds.ols <- predict(ols, new = test)
rmse(test$Grad.Rate, preds.ols)
rmse(train$Grad.Rate, preds.ols)
```

```{r}
preds.train <- predict(model.1se, newx = train_x)
preds.train1 <- predict(model1.1se, newx = train_x)
train.rmse <- rmse(train_y, preds.train)
train1.rmse <- rmse(train_y, preds.train1)
```

```{r}
preds.test<- predict(model.1se, newx = test_x)
test.rmse <- rmse(test_y, preds.test)

preds.test1<- predict(model1.1se, newx = test_x)
test1.rmse <- rmse(test_y, preds.test1)

train1.rmse
train.rmse
test.rmse
test1.rmse


```
